---
title: "p8105_hw3_kz2334"
author: "Kangkang Zhang"
date: "10/8/2018"
output: github_document
---

##Problem 1 


Load required packages.
```{r 1.1}
library(tidyverse)
devtools::install_github("p8105/p8105.datasets")
library(p8105.datasets)
```

---

Load the BRFSS data and do some data cleaning.
```{r 1.2}
data(brfss_smart2010)
brfss_data = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health", response %in% c("Excellent", "Very good", 
                                                  "Good", "Fair", "Poor")) %>% 
  mutate(response = factor(response, ordered = TRUE, levels = 
                          c("Poor", "Fair", "Good", "Very good", "Excellent")))
```

---

Find out which states were observed at 7 locations in 2002.
```{r 1.3}
brfss_data %>% 
  distinct(locationdesc, locationabbr, year) %>% 
  count(locationabbr, year) %>% 
  filter(year == 2002, n == 7)
```

CT, FL NC were observed at 7 locations in 2002. It shows that the program had great attention to those three states in 2002.

---
Make a “spaghetti plot” that shows the number of locations in each state from 2002 to 2010.
```{r 1.4}
brfss_data %>% 
  distinct(locationdesc, locationabbr, year) %>% 
  count(locationabbr, year) %>% 
ggplot(aes(y = n, x = year, color = locationabbr)) + 
  geom_line() + 
  labs(
    title = "spaghetti plot",
    x = "Year",
    y = "Number of Locations"
  ) +
  theme_bw() + 
  viridis::scale_color_viridis(
    name = "State", 
    discrete = TRUE
  ) 
  
```

We can see that the numbers of locations in majority of states fluctuated within [0, 10] from 2002 to 2010. There was a state with number less than 10 in 2006, then rapidly grow up to larger than 40 in 2007. 
